
\section{Introduction} \label{sec:introc1}

Financial literacy is at the center of the education debate in various developed and developing countries. The Organization for Economic Cooperation and Development (OECD) advocates that financial education should be integrated into basic school curricula \citep{Lusardi2014}.\footnote{The OECD defines financial literacy as a combination of awareness, knowledge, skills, attitudes, and behaviors necessary to make sound financial decisions and ultimately achieve individual financial well-being. In this study, we use financial education and financial literacy interchangeably.} As individuals are being increasingly exposed to financial decisions at a young age, the sooner they learn to incorporate the costs of their choices in their decision-making process the better. Good financial habits at an early age are likely to benefit schooling, employment, and the standard of living through adulthood (\cite{Bruhn2016}; \cite{bruhn2022long}). In fact, developed and developing countries worldwide have been implementing financial education interventions in schools aiming to develop financial skills that could be conducive to better decision-making throughout individuals' life-cycle (\cite{bover2018impact}, \cite{frisancho2020impact}, \cite{gill2019effects}, \cite{Luhrmann2018}), \cite{Bruhn2016}, \cite{bhattacharya2016effectiveness}, \cite{de2021effect}, \cite{berry2018impact}, \cite{batty2020experiential}).

Financial education may improve the quality of intertemporal decision-making by increasing
individuals’ knowledge about saving and borrowing instruments (\cite{Luhrmann2018}) as subjects with higher financial knowledge are more likely to make patient inter-temporal choices (\cite{oberrauch2022cognitive}). Also, financial literacy can be a factor in avoiding financial risk and taking advantage of economic opportunities (\cite{Miller2015}). In a recent meta-analysis, \cite{kaiser2022financial} find that financial education interventions have, on average, a significant impact on financial knowledge (0.15-0.2 of a standard deviation) and on financial behaviors, such as borrowing, saving and investing, budgeting and planning, insurance, and remittances  (0.06–0.1 sd).\footnote{The literature review includes 76 randomized experiments.} 

Effective financial education programs that impact children's socio-emotional skills such as patience and self-control could potentially improve outcomes other than the financial-related ones. The more impatient children and adolescents are less likely to invest in human capital and seem more likely to consume alcohol, develop obesity, get pregnant during adolescence, and commit a crime (\citealt{castillo2011today}; \citealt{moffitt2011gradient}; \citealt{sutter2013impatience}; \citealt{golsteyn2014adolescent}; \cite{kautz2014fostering}). These same socioemotional skills can influence labor market prospects in future life by affecting participation in the labor market, the probability of holding a job, school decisions, and a range of educational outcomes (\cite{acosta2018role}; \cite{santos2021can}; \cite{alan2019ever}; \cite{kautz2014fostering}).
From a theoretical standpoint, programs aimed at developing personality traits as such should be prioritized in childhood in the face of the expected higher returns per dollar invested (\citealt{heckman2006effects}; \citealt{cunha2007technology}). 

There is increasing evidence that financial education programs targeting high school students are effective (\cite{Bruhn2016}, \cite{gill2019effects}; \cite{frisancho2022school}). The available literature also suggests that in-school programs not only improve students' financial knowledge but also change their behavior (\cite{bover2018impact}, \cite{Luhrmann2018}, \cite{frisancho2022school}, \cite{jamison2014financial}, \cite{Bruhn2016}, \cite{stoddard2020effects}, \cite{urban2020effects}, and \cite{harvey2019impact}). Recent evidence suggests that these programs can have long-lasting effects on students' human capital accumulation and on financial decisions and entrepreneurship (\cite{frisancho2020impact};\cite{frisancho2022school}; \cite{bruhn2022long}). 

There is also growing interest and evidence on the positive impacts of financial literacy interventions on elementary and middle school grades. 
\cite{bhattacharya2016effectiveness} find evidence of a 12 percentage points increase in the financial knowledge of eighth-graders in the United States. \cite{de2021effect} suggest a 0.46 sd increase in the financial knowledge as well as better financial behavior of eighth and ninth-graders in an experiment in Belgium. \cite{batty2020experiential} finds evidence of an increase in the financial knowledge of fourth-graders equivalent to 0.25 of a standard deviation in a pilot run in 20 schools in the United States. In an RCT with Turkish elementary schools, \cite{alan2018fostering} find that the program persistently reduces the incidence of behavioral problems in schools and increases students' patience. \cite{berry2018impact}, on the other hand, do not find impacts of a financial literacy intervention implemented in Ghanaian elementary and junior high schools.   

In this paper, we use a cluster randomized controlled trial to investigate the impacts of a financial literacy pilot implemented in 2015 in elementary and middle school grades in two Brazilian municipalities (Manaus and Joinville). We randomly assigned 101 municipal schools to receive the program, and 100 to control. The intervention aimed to help young students become more conscious of the trade-offs embedded in their daily decisions and more forward-looking. The evaluation focused on grades three, five, seven, and nine for reasons discussed later.  

We find positive effects on students' financial knowledge of 0.07 of a standard deviation, on average. The effects are driven by middle school grades (0.1 SD) and are consistent with a recent summary of the financial education literature (\cite{kaiser2022financial}). We also find suggestive evidence that the program helped students become more conscious about their consumption habits - i.e., more patient and risk-averse. We find suggestive evidence of changes in some contemporary behavioral outcomes such as the use of piggy banks, suggesting some linkages between actual change in behavior in the short-term. However, we do not detect impacts on both contemporary learning outcomes and human capital accumulation. We use a mediation causal analysis to investigate the potential causal link between gains in financial knowledge and changes in behavioral outcomes. Our results do not point to a compelling causal linkage between formal knowledge and change in behavior, a result consistent with \cite{de2021effect}.

Our paper has four main contributions. First, we document the positive impacts on financial proficiency, attitudes, and behavioral outcomes of a large-scale pilot intervention implemented in elementary and middle school grades.  Second, we show that financial education programs can potentially impact different skills according to students' life cycles. Third, we test the overlooked hypothesis that increasing financial proficiency is a necessary condition to change individuals' behavior. Finally, we use student-level data from the Brazilian national standardized exams to assess the program's impacts on students' contemporary and longer-term human capital. 


Apart from this introduction, this paper is organized as follows. Section \ref{sec:pilot} describes the pilot program. Section \ref{sec:data_c} describes the data collection. Section \ref{sec:implementation} describes the program's implementation. Section \ref{sec:identification} introduces the empirical strategy, whereas section \ref{sec:results} discusses the results. We then conclude.

