\section {Financial Education Pilot in Elementary and Middle Schools  } \label{sec:pilot}

The(\textit{Estratégia Nacional de Educação Financeira – ENEF}) was established by the Brazilian Federal Government in 2010 to promote financial education and empower individuals to make autonomous financial decisions. Between August 2010 and December 2011, ENEF implemented a pilot financial education program in public high schools across six states—São Paulo, Rio de Janeiro, Ceará, Tocantins, Minas Gerais and the Federal District. To assess the program’s impact, a randomized controlled trial (RCT) was conducted, involving 891 schools and approximately 20,000 students. The evaluation revealed significant improvements in financial knowledge, saving intentions, and financial autonomy, as well as increased student participation in household financial decisions \citep{Bruhn2016}.

Building on the success of the high school pilot, ENEF expanded its initiative to primary public schools. In 2014, new educational materials were developed, and in 2015, the program was piloted in municipal schools in Joinville and Manaus, with their respective Municipal Secretariats of Education voluntarily joining the initiative. Notably, these two cities represent distinct socioeconomic contexts: Manaus, the capital of Amazonas, is located in northern Brazil, while Joinville, a municipality in the southern state of Santa Catarina, is in a more affluent region. By 2016, Joinville exhibited higher per capita income (USD 14,000 vs. 11,000), a superior Municipal Human Development Index (MHDI) (80.9 vs. 73.7), higher school attendance rates among 6- to 14-year-olds (97\% vs. 94\%), and stronger performance on the Brazilian Education Development Index (IDEB) in both primary and secondary education. While these differences may introduce regional variations in program implementation, they also carry important external validity implications, reinforcing the broader relevance of the pilot’s findings.

The pilot for elementary and middle school students consisted of one module per grade (1st to 9th), forming a teaching material designed to be integrated cross-sectionally into the existing curriculum. Teachers of all subjects are encouraged to incorporate the textbook into their lessons, and the teaching materials recommend educators in developing an annual plan for its implementation. While the modules are independent, they are designed to complement each other over time, reinforcing six comun learning objectives: (i) educating for citizenship, (ii) promoting ethical, conscious, and responsible consumption and saving, (iii) providing concepts and tools for autonomous decision-making based on attitude change, (iv) training multipliers to disseminate knowledge, (v) fostering skills for short-, medium-, and long-term planning, and (vi) cultivating a culture of prevention. 


Four different pedagogical approaches. The first, for students from first to fourth graders,
the second one for fifth and sixth graders, the third one for seventh and eighth graders, and the fourth one for ninth graders

one for students in the 1\textsuperscript{st} to 4\textsuperscript{th} elementary school grades, one for students in the 5\textsuperscript{th} and 6\textsuperscript{th} grades, one for students in the 7\textsuperscript{th} and 8\textsuperscript{th} grades, and an exclusive module for students in 9\textsuperscript{th} grade. As with the material developed for high school students in Brazil, these 

The teaching material emphasizes the role of financial education in students’ everyday life by developing financial concepts based on spatial and temporal dimensions. The spatial dimension is built upon notions at individual, local, regional, national, and global levels and upon the effects of individual actions on the social context. The temporal dimension, however, deals with the intertemporal nature of financial decisions. 


 Table \ref{tab:descriptors} in Appendix shows the descriptors and skills associated with each grade.








ENEF prepared four financial education textbooks for elementary and middle school students. The first one is for first to fourth graders, the second one for fifth and sixth graders, the third one for seventh and eighth graders, and the fourth one for ninth graders. Each textbook is tailored to be adequate for students' age groups and grades. \autoref{tab:descriptors} shows the skills the intervention aimed to develop. The curriculum is conceived to introduce the role of financial education in students’ daily lives by developing financial concepts such as savings, consumption, and waste; making students able to identify situations related to financially responsible attitudes, estimate the budget necessary for financial projects, and find relevant information in the press media for decision-making in finance. With the development of these skills, the pilot is expected to increase students' preference for future prospects and decrease present bias (\cite{becker1997endogenous}; \cite{Luhrmann2018}). The textbooks' content is then transversally introduced into each grade's standard curriculum.  

\subsection{Sample Selection}

Due to budgetary limitations, we agreed with the implementing partner to identify approximately 200 schools that could participate in the pilot. We leverage existing administrative data to define the sampling frame of the experiment. According to the 2015 Census of Education, Joinville has 72 schools that offer elementary or middle school grades. We include them all in the experiment sampling frame.\footnote{Public schools that the municipal government manages.} In Manaus has 302 schools and we randomly draw 129 to be part of the experiment sampling frame (\autoref{tab:sample}).\footnote{The total of 302 schools does not include 53 schools in riverside communities in Manaus.} For budgetary reasons, the pilot could not include all K-9 grades. In agreement with ENEF, we select grades 3, 5, 7, and 9 to generate evidence on the effects of each of the four textbooks developed for the pilot. This strategy allows us to assess whether the effect of the program varies by school grade and students' like-cycle.\footnote{We are not testing the impact of the same material in different school grades. Instead, we are testing the impacts of school materials tailored to each specific grade.}  Within each treated school, not all the classes offering third, fifth, seventh, and ninth grade participate in the financial literacy intervention. At the beginning of the 2015 school year, the school principals from both treatment and control schools picked only one class of each grade to be included in the pilot. This strategy is adopted to minimize costs with data collection. 

In the end, the pilot included 201 schools, with 101 randomly assigned to the treatment (9,204 students), and 100 assigned to the control (9,353 students).To account for differences in both municipalities' socioeconomic standards and schools' infrastructure, we stratify the sample by the municipality and three school types: schools that only offer elementary school grades (first to fifth grades in Brazil), schools that only offer middle school grades (sixth to ninth grades in Brazil), and schools that offer all elementary school grades, totaling six strata (\autoref{tab:sample}).

********************************************************
Por exemplo, em Língua Portuguesa, o livro traz oportunidades interessantes para se trabalhar 
a competência de leitura dos alunos, como também o conceito de autoria, pedindo aos alunos que criem alternativas diferentes para as histórias. Na disciplina de Matemática, o livro possibilita que o tra
balho privilegie o caráter construtivo do conhecimento matemático permitindo ao aluno compreender e transformar sua realidade, deenvolver a capacidade de observar, estabelecer relações, argumentar 
matematicamente e a relacionar observações do mundo real com representações (esquemas, tabelas, fi guras, escritas numéricas) e relacionar estas representações com princípios e conceitos matemáticos. 

1 ao 4 ano
 O programa de Educação Financeira se apresenta como uma estratégia para contribuir com os resultados dos alunos brasileiros em exames de aferição da aprendizagem em disciplinas básicas como Português e Matemática. Entretanto, é necessário contar com as contribuições de outras áreas do conhecimento, embora uma visibilidade maior permaneça voltada para Português e Matemática. E não se trata de inserir pedaços de aulas específicas das disciplinas escolares nas páginas dos livros de Educação Financeira, mas de tomar tais disciplinas como referenciais e como ferramentas para explorar os conhecimentos da vida financeira.

O material didático de Educação Financeira, nos quatro anos iniciais, compõe-se de um volume para o aluno, contendo roteiros de trabalho para cada um dos quatro projetos de cada ano escolar, e de um livro para o professor que apresenta e discute orientações pedagógicas para trabalhar os conteúdos em sala de aula. Os livros do aluno acompanham a estética de livros de literatura infantil. Os dois primeiros volumes atendem a idades mais tenras e, por isso, apresentam um traçado que remete à fantasia, economizam nos detalhes e utilizam cores mais suaves. Os dois volumes seguintes têm ilustrações com mais de
talhes e de cunho realista, assim como utiliza cores mais fortes, em atendimento às caraterísticas da faixa etária. 

Os projetos de trabalho seguem um ciclo que se defl agra com uma pergunta inicial e se encerra com respostas encontradas pela ação das ferramentas do pen
samento sobre os conhecimentos de vida real e de Educação Financeira.

O fato de o livro ser de Educação Financeira e não de uma disciplina escolar específi ca, faz com que informações de natureza histórica, geográfi ca, artística etc. sejam apresentadas apenas com o objetivo de contribuir para a construção de conceitos e de comportamentos fi nanceiros. Daí se explica o fato 
de não receberem tratamento minucioso nem indicação de referência para 
aprofundamento. 
Diferentemente disso, para as áreas de Matemática e de Língua Portuguesa  são apresentadas indicações bibliográfi cas, que poderão contribuir com o 
aprendizado dos educandos. O Projeto Pedagógico do Programa para os 
anos iniciais se encontra explicitamente comprometido com a melhoria dos 
resultados dos educandos nessas duas áreas do conhecimento.

4 projetos em 4 eixos temáticos: Produção e Consumo, Orgamnização, Cuidados, Planejamento