\section{Data and Measurement} \label{sec:data_c}

To evaluate the impact of the financial education pilot, we use a combination of survey and administrative data to measure a broad set of outcomes over different time horizons. Survey data collected at the end of the 2015 school year capture short-term effects on students’ financial proficiency, hypothetical attitudes toward consumption and saving, and self-reported behaviors. The survey data were collected right after the program implementation finished, in line with other financial education programs targeting elementary and middle school students (\cite{de2021effect}, \cite{alan2018fostering}, \cite{berry2018impact}). Administrative records provide medium- and long-term outcomes, including grade progression, standardized test scores, and dropout rates, observed one to three years after the intervention.


\subsection{Survey Data} \label{subsec:question}

The Center for Public Policy and Education Evaluation (CAEd), a Brazilian survey firm specializing in standardized assessments and data collection, was responsible for implementing the survey, with students and teachers participating in the pilot. Data collection occurred during the final school days of the 2015 academic year, when CAEd distributed printed questionnaires directly in classrooms. Despite being conducted at the end of the school year—a period when some students may already be on vacation or have dropped out—student participation rates remained high, averaging approximately 80\%, comparable to regular school-day attendance. Furthermore, as shown in \autoref{tab:implementation}, there is no significant difference between participation rates of treatment and control groups across all grades.

%\footnote{\autoref{tab:implementation}  shows the students' participation rates. One may wonder whether there could be less student participation with the survey data being collected at the end of the school year as some students might have already gone on vacation or dropped out. On the one hand, the participation rate was close to the one observed on a regular school day. On the other hand, it is unclear if the absent students are the best ones, who had already succeeded in being promoted to the next grade even without final tests, or the worst students who had already dropped out or did not achieve the expected performance for grade promotion.}

Students completed three paper-based questionnaires during a two-hour session: (i) a financial literacy test, (ii) a questionnaire on hypothetical consumption and savings attitudes and self-reported behaviors, and (iii) a socioeconomic questionnaire. The content of each instrument was tailored to the students’ grade level and cognitive development.

The financial literacy test evaluates students’ understanding of concepts introduced in the financial education textbooks through multiple-choice questions aligned with grade-appropriate skills (\autoref{tab:descriptors}). The test includes items that assess students’ ability to interpret everyday financial information, such as utility bills or spending records, and to identify the purpose of documents like receipts, checks, and invoices. It also covers savings, expenses, consumption, waste, risk, return, financial planning, and investment. Additional questions assess students’ ability to estimate values for simple financial projects, distinguish between paid and unpaid work, interpret financial graphs and tables, and extract relevant financial information from media texts. Not all skills are assessed in all grades; the test content is adapted to the cognitive level of each group, with some skills appearing only in earlier or later grades.\footnote{For third-grade students, the financial literacy test was read aloud by trained enumerators to ensure accessibility, given that students at this age are still developing reading and writing skills.}

CAEd used student responses to construct a financial proficiency index using Item Response Theory (IRT), which accounts for the difficulty of each question and generates a standardized score comparable across grades. This measure allows us to assess the impact of the intervention in terms of standard deviations, enabling consistent comparisons within and across student groups.

The second questionnaire assesses students’ financial attitudes and self-reported behaviors related to saving and consumption. Attitudinal items present hypothetical scenarios (e.g., “I plan before spending,” “I see no problem in owing money”), with responses captured using a four-point Likert scale (“Totally agree” to “Totally disagree”), where 4 reflects the most forward-looking or prudent financial attitude (\autoref{tab:cs_index}). Based on these responses, CAEd constructed two standardized indices: a consumption index, reflecting students’ tendency toward planned versus impulsive spending (self-control), and a savings index, reflecting attitudes related to risk aversion and the importance of saving for the future. These indices are derived using principal component analysis and follow the methodology of psychology-based indices used in similar studies (e.g., \cite{Bruhn2016}). Importantly, these measures reflect students’ attitudes, not their underlying preferences or actual financial behavior. This second questionnaire also collects self-reported behavioral outcomes, such as whether students have a piggy bank, receive an allowance, use a credit card, or talk about money with friends or family. These variables are analyzed separately as self-reported behaviors.
 
The last questionnaire asks about students' socioeconomic background, such as the educational attainment of their mothers. Whether they benefit from the national conditional cash transfer program, \textit{Bolsa Família}. Fifth, seventh, and ninth graders answered all the questionnaires in school.  Third-graders answer the financial knowledge questionnaire in school with the help of an enumerator, who reads the questions. Since third-graders are still acquiring reading and writing abilities, this strategy is intended to make the test easier for all students, regardless of their reading and writing skills. Still, for third-graders, the socioeconomic questionnaire and one of the students' attitudes towards savings and consumption were sent to their homes so their legal guardians could fill them out and return them the following day. Legal guardians were supposed to answer it based on how he/she believed his/her child would behave in a given situation. Schools prepared the legal guardians to answer these questionnaires by sending them a note before the test day.  \autoref{tab:implementation} shows that over three-quarters of the parents/guardians returned the questionnaire, and that there is no statistically significant difference between the percentage of guardians that completed the questionnaire for treated and control students <WE DID NOT CONSIDER SAVINGS AND CONSUMPTION FOR THIRD GRADERS). 

As expected, \autoref{tab:studentscar} shows no significant differences between the socioeconomic characteristics of treated and control students. The teachers' questionnaire investigates their wage level, teaching experience, and issues related to the incorporation of financial literacy textbooks in the curriculum, among other questions. Overall, there are no significant differences between teachers in treatment and control groups (\autoref{tab:teacherscar}). 

Overall, our analysis has eight dependent variables from the survey data collected in the field. Three standardized indices have been developed by CAEd, which are (i) financial proficiency, (ii) consumption (a proxy for patience and self-control), and (iii) savings (a proxy for risk aversion). The other five are behavioral ones such as (iv) a dummy indicating whether a student talks to parents about financial subjects, (v) a dummy indicating whether a student talks to friends about the same subject, (vi) students' use piggy banks (a proxy for saving), (vii) students' access to financial services, and (viii) a dummy indicating whether a student receives an allowance from their parents.  


\subsection{Administrative Data}

Due to the limited budget and the richness of publicly available administrative data, we did not collect baseline survey data. Prior to the program's implementation, we used administrative data at the school level to check for a balance between treatment and control schools. 

Since 1995, all private and public K-12 schools have participated in the annual Education Census. The Census is implemented by the National Institute of Educational Studies and Research (INEP), a research agency under the Brazilian Ministry of Education. The Census collects information on school facilities, school infrastructure, social services, teachers, and students.\footnote{The data is collected at the student level, allowing us to track their sex, ethnicity, age, grade level, instruction time per day, class size, and student outcomes, such as grade promotion, retention, and dropout status. In 2005, INEP introduced the national standardized exam known as \textit{Prova Brasil} that applies proficiency assessments in reading and math to all public schools.\footnote{Schools with at least 20 students enrolled in each evaluated grade. Students take the test at the end of the school year (between October and November). The exam is applied every two years to fifth and ninth-graders, and students in the last high school grade. Since 1998, INEP has annually applied the National High School Proficiency Test (ENEM).\footnote{ENEM stands for \textit{Exame Nacional do Ensino Medio}. The exam aims to assess students' learning levels in reading, math, human sciences, and natural sciences. \footnote {ENEM is not mandatory, and the students are responsible for registering for the exam. ENEM participants are the students enrolled in high school, mostly in the last grade, and those who have already finished this level of education. Since 2009, the exam has been used by several Brazilian universities as one of their criteria for selecting their students. Consequently, the number of participants significantly increased, from 115,000 in 1998 to 8.5 million in 2015.} 

We merged the pilot data with the surveys above, all of which are public at the student level.  \autoref{tab:pret_balancing} shows that treated and control schools do not have significant differences in average students' proficiency in reading and math, grade promotion, dropout, retention rates, and age-grade distortion.\footnote{Unidentified microdata at the student level is also publicly available and identified data only upon researcher request to access INEP's facilities. We could not request access to students' identified microdata by the beginning of 2015 when the pilot was implemented, and then check the balance between treated and control students. First, we did not have the list of pilot students; second, the request usually takes months to be processed, and it was very restrictive then.} However, as presented in \autoref{sec:pilot}, principals' choice of which classes would receive the intervention or serve as a control is not necessarily random. In this sense, one may wonder whether school principals from treatment and control schools face different incentives to select the classes they did. In treated schools, principals could either select low-performing classes, hoping to help their students with the intervention, or high-performing classes if they believe these students would benefit most from the program. In control schools, principals could select low performers to show that their schools should also receive the treatment in the future, or high performer classes if they intend to show that their schools do not need the program. In this sense, using administrative data at the school level to check the pre-intervention balance of students' performance has one caveat, since not all the students in these schools are included in the pilot.

To enrich our analysis and compare students instead of schools, in 2019, we requested INEP to merge our pilot data with their administrative datasets at the student level 
level.\footnote{This task has to be performed by the agency since researchers cannot access identified administrative data at the student level. The INEP technical team performs the merge using students' names and the schools they are enrolled in. \autoref{tab:balance_adm1} and \autoref{tab:balance_adm2} show no significant differences in the percentage of treated and control students found by the INEP technical team in its administrative datasets. All the analyses with administrative data at the student level are run in the cold room of INEP facilities.}  \autoref{tab:inep_data} shows the variables that we were able to collect and we use them for two main tasks. First, the merged dataset allows us to investigate the impact of the financial literacy pilot on student outcomes only available in Education Census, \textit{Prova Brasil}, and \textit{ENEM} such as proficiency in reading and math, grade-promotion, retention and dropout (\autoref{tab:inep_data}).\footnote{\autoref{tab:inep_data} shows the administrative data that we use to assess the effects of the financial literacy pilot on (i) retention and dropout rates in the year of the pilot or up to three years later (between 2015-2018); on (ii) standardized proficiency test in math and Portuguese in the year of the pilot (for those enrolled in fifth and ninth grades in 2015) and two years after (those enrolled in third and seventh grades in 2015) - for those who did not repeat between third and fifth grades and between seven and ninth grades, respectively; and (iii) ENEM scores and high school completion rates for those enrolled in the ninth grade in the year of the pilot.\footnote{For those who did not repeat between the ninth grade and the last year of high school.} }

Second, for seventh and ninth graders included in the pilot, we compare their pre-intervention performance in reading and math, since these students did the proficiency exam when enrolled in fifth grade.\footnote{For seventh graders who did not repeat any grade between the fifth and seventh grades, their fifth-grade proficiency is available in 2013. For ninth-graders who did not repeat any grade between the fifth and ninth grades, their fifth-grade proficiency is available in 2011.} \autoref{tab:balance_adm2} shows that, for seventh-grade classes, there are no significant differences between treatment and control classes in pre-treatment students' performance. Nonetheless, for ninth-grade classes, the data show that students in control schools had a significantly higher pre-treatment performance in reading, which might indicate that the principals from control schools selected high-performing classes to be in the pilot evaluation. If this were the case, and if higher reading scores are correlated with better financial proficiency skills, our results would be biased downwards. \footnote{No standardized proficiency test is applied by INEP before the fifth grade. Therefore, for third and fifth graders, we do not have any administrative data on their pre-treatment performance in reading and math on standardized national exams. Hence, we could not compare students' proficiency between treatment and control classes for these grades. } 









